# Data transformation

We typecasted the date and time columns to the required format. When plotting Borough, Zip Code and Crash time summary statistics, we dropped all instances where the values were not available and grouped by the appropriate variable. Depending on the graph requirements, we took subsets of the data using filtering. When plotting the Choropleth maps, we obtained the geojson data corresponding to NYC Boroughs and Zip Codes, and merged it with our transformed data.

```{r}
library(dplyr)
library(Lock5withR)
library(tidyverse)
library(patchwork)
library(ggplot2)
library(ggnewscale)
library(hms)
library(lubridate)
library(scales)
library(tigris)
library(cartography)
library(sf)
library(geojsonio)
library(broom)
```


```{r}
crash <- read_csv('Motor_Vehicle_Collisions_Crashes.csv',na = c("", "NA", " "))
names(crash) <- make.names(names(crash), unique=TRUE)
crash<-crash%>% drop_na(CRASH.TIME)
crash$CRASH.TIME <- as.hms(crash$CRASH.TIME)
crash<-crash[ order(crash$CRASH.TIME , decreasing = FALSE ),]
crash$CRASH.TIME.hr.range <- hour(crash$CRASH.TIME)
head(crash)
```

```{r}
a<-crash%>%group_by(CRASH.TIME.hr.range)%>%count()
b<-crash%>%filter(NUMBER.OF.PERSONS.KILLED>0)%>%group_by(CRASH.TIME.hr.range)%>%count()
b$n<-(b$n/a$n)*100
names(b)<- c('Hours', 'Percentage of fatal accidents')
head(b)
```

```{r}
crash$CRASH.DATE <- as.Date(crash$CRASH.DATE , format = "%m/%d/%Y")
monthly<-crash %>%
mutate(month = format(CRASH.DATE, "%m")) %>%
group_by(month) %>%
count()
monthly$month=seq(1,12,1)
monthly$month <- as.Date(paste0("2018-", monthly$month, "-1"))
monthly$month <- months(as.Date(monthly$month))
names(monthly)<- c('Months', 'Accident.Count')
monthly_f<-crash %>%filter(NUMBER.OF.PERSONS.KILLED>0)%>%
mutate(month = format(CRASH.DATE, "%m")) %>%
group_by(month) %>%
count()
monthly_f$month=seq(1,12,1)
monthly_f$month <- as.Date(paste0("2018-", monthly_f$month, "-1"))
names(monthly_f)<- c('Months', 'Fatal.Accident.Count')
monthly$Fatal.Accident.Count<-monthly_f$Fatal.Accident.Count
head(monthly)
```
```{r}
daily<-crash %>%
mutate(month = format(CRASH.DATE, "%m"), day = format(CRASH.DATE, "%d")) %>%
group_by(month, day) %>%count()
daily$date=seq(as.Date("2012-01-01"), as.Date("2012-12-31"),1)
yo<-daily[c('month','day','n')]
names(yo)<-c('Month', 'Day', 'Accident Count')
head(yo)
```

```{r}
bor<-crash%>% drop_na(BOROUGH)%>% group_by(BOROUGH)%>%count()
bor$BOROUGH<-str_to_title(bor$BOROUGH)
bw<-crash%>%filter(BOROUGH =='BROOKLYN' | BOROUGH =='STATEN ISLAND')%>%group_by(BOROUGH,CRASH.TIME.hr.range)%>%count()
bw$n[bw$BOROUGH == 'BROOKLYN'] <- -bw$n
head(bw)
```

```{r}
contour<-crash%>%drop_na(LATITUDE,LONGITUDE)%>%filter(LATITUDE>0)%>%filter(LONGITUDE>(-74.5))%>%filter( LONGITUDE<(-50))%>%filter(LATITUDE<41)%>%drop_na(BOROUGH)%>% 
  group_by(LATITUDE, LONGITUDE, BOROUGH) %>%
  summarise(Total = n())
head(contour)
```

```{r}
crash$year<-format(crash$CRASH.DATE, "%Y")
yearly_2016<-crash%>%drop_na(ZIP.CODE)%>%filter(year<2017)%>%group_by(ZIP.CODE,year) %>%count()%>%
pivot_wider(names_from = year,
values_from = n)
yearly_2020<-crash%>%drop_na(ZIP.CODE)%>%filter(year>2016)%>%group_by(ZIP.CODE,year) %>%count()%>%
pivot_wider(names_from = year,
values_from = n)
yearly_2016<-yearly_2016 %>% mutate(mean_first = rowMeans(across(where(is.numeric)),na.rm=TRUE))%>%select(ZIP.CODE,mean_first)
yearly_2020<-yearly_2020 %>% mutate(mean_second = rowMeans(across(where(is.numeric)),na.rm=TRUE))%>%select(ZIP.CODE,mean_second)
yearly<-merge(x=yearly_2016,y=yearly_2020,by="ZIP.CODE")%>%mutate(Difference=mean_second-mean_first)
yearly<-yearly[order(yearly$Difference),]
yearly<-rbind(head(yearly,n=5),tail(yearly,n=5))
head(yearly)
```

```{r}
data <- read.csv(file = 'Motor_Vehicle_Collisions_Crashes.csv', na.strings = c("", "NA", " "))
cols = c('BOROUGH','ZIP.CODE','LATITUDE','LONGITUDE','LOCATION')
data2 <- data[!is.na(data$BOROUGH), 0:18]
data2 <- data2[-c(8:10)]

data2$TOT.KILLED <- data2$NUMBER.OF.PEDESTRIANS.KILLED + data2$NUMBER.OF.CYCLIST.KILLED + data2$NUMBER.OF.MOTORIST.KILLED

data2$TOT.INJURED <- data2$NUMBER.OF.PEDESTRIANS.INJURED + data2$NUMBER.OF.CYCLIST.INJURED + data2$NUMBER.OF.MOTORIST.INJURED

data2$NUMBER.OF.PERSONS.KILLED <- ifelse(is.na(data2$NUMBER.OF.PERSONS.KILLED),data2$TOT.KILLED,data2$NUMBER.OF.PERSONS.KILLED)

data2$NUMBER.OF.PERSONS.INJURED <- ifelse(is.na(data2$NUMBER.OF.PERSONS.INJURED),data2$TOT.INJURED,data2$NUMBER.OF.PERSONS.INJURED)

head(data2)
```
```{r}

x <- data2
x$YEAR <- format(as.POSIXct(x$CRASH.DATE, format = "%m/%d/%Y"),format = '%Y')
t <- x %>% group_by(YEAR,BOROUGH) %>% 
    summarise(NUMBER.OF.PERSONS.KILLED = sum(NUMBER.OF.PERSONS.KILLED),.groups = 'drop')
head(t)
```

```{r}
xB <- data2[data2$BOROUGH=='BROOKLYN',]
xB_new <- xB %>% group_by(ZIP.CODE) %>% 
    summarise(NUMBER.OF.PERSONS.TOT.HURT = sum(NUMBER.OF.PERSONS.KILLED) + sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')
xB_new <- xB_new[order(-xB_new$NUMBER.OF.PERSONS.TOT.HURT),]
xB_new <- xB_new %>% mutate(ZIP.CODE = as.character(ZIP.CODE))
head(xB_new)
```


```{r}
data2$YEAR <- format(as.POSIXct(data2$CRASH.DATE, format = "%m/%d/%Y"),format = '%Y')

incidents_tma <- data2 %>% group_by(YEAR, BOROUGH) %>% summarise(Total.Hurt = sum(NUMBER.OF.PERSONS.INJURED) + sum(NUMBER.OF.PERSONS.KILLED),.groups = 'drop')
head(incidents_tma)
incidents_tma2 <- data2 %>% group_by(YEAR, BOROUGH) %>% summarise(Total.Injured = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')
head(incidents_tma2)
incidents_tma3 <- data2 %>% group_by(YEAR, BOROUGH) %>% summarise(Total.Killed = sum(NUMBER.OF.PERSONS.KILLED),.groups = 'drop')
head(incidents_tma3)
```

```{r}
data2 <- data[!is.na(data$BOROUGH), ]
data2 <- data2[-c(8:10)]

data2$TOT.KILLED <- data2$NUMBER.OF.PEDESTRIANS.KILLED + data2$NUMBER.OF.CYCLIST.KILLED + data2$NUMBER.OF.MOTORIST.KILLED

data2$TOT.INJURED <- data2$NUMBER.OF.PEDESTRIANS.INJURED + data2$NUMBER.OF.CYCLIST.INJURED + data2$NUMBER.OF.MOTORIST.INJURED

data2$NUMBER.OF.PERSONS.KILLED <- ifelse(is.na(data2$NUMBER.OF.PERSONS.KILLED),data2$TOT.KILLED,data2$NUMBER.OF.PERSONS.KILLED)

data2$NUMBER.OF.PERSONS.INJURED <- ifelse(is.na(data2$NUMBER.OF.PERSONS.INJURED),data2$TOT.INJURED,data2$NUMBER.OF.PERSONS.INJURED)

data2$CONTRIBUTING.FACTOR.VEHICLE.1[data2$CONTRIBUTING.FACTOR.VEHICLE.1=='Illnes'] <- 'Illness'
data2$CONTRIBUTING.FACTOR.VEHICLE.2[data2$CONTRIBUTING.FACTOR.VEHICLE.2=='Illnes'] <- 'Illness'
data2$CONTRIBUTING.FACTOR.VEHICLE.3[data2$CONTRIBUTING.FACTOR.VEHICLE.3=='Illnes'] <- 'Illness'
data2$CONTRIBUTING.FACTOR.VEHICLE.4[data2$CONTRIBUTING.FACTOR.VEHICLE.4=='Illnes'] <- 'Illness'
data2$CONTRIBUTING.FACTOR.VEHICLE.5[data2$CONTRIBUTING.FACTOR.VEHICLE.5=='Illnes'] <- 'Illness'

data2$CONTRIBUTING.FACTOR.VEHICLE.1 <- tolower(data2$CONTRIBUTING.FACTOR.VEHICLE.1)
data2$CONTRIBUTING.FACTOR.VEHICLE.2 <- tolower(data2$CONTRIBUTING.FACTOR.VEHICLE.2)
data2$CONTRIBUTING.FACTOR.VEHICLE.3 <- tolower(data2$CONTRIBUTING.FACTOR.VEHICLE.3)
data2$CONTRIBUTING.FACTOR.VEHICLE.4 <- tolower(data2$CONTRIBUTING.FACTOR.VEHICLE.4)
data2$CONTRIBUTING.FACTOR.VEHICLE.5 <- tolower(data2$CONTRIBUTING.FACTOR.VEHICLE.5)

data2$VEHICLE.TYPE.CODE.1 <- tolower(data2$VEHICLE.TYPE.CODE.1)
data2$VEHICLE.TYPE.CODE.1[data2$VEHICLE.TYPE.CODE.1=='station wagon/sport utility vehicle'] <- 'sport utility / station wagon'
data2$VEHICLE.TYPE.CODE.2 <- tolower(data2$VEHICLE.TYPE.CODE.2)
data2$VEHICLE.TYPE.CODE.2[data2$VEHICLE.TYPE.CODE.2=='station wagon/sport utility vehicle'] <- 'sport utility / station wagon'
data2$VEHICLE.TYPE.CODE.3 <- tolower(data2$VEHICLE.TYPE.CODE.3)
data2$VEHICLE.TYPE.CODE.3[data2$VEHICLE.TYPE.CODE.3=='station wagon/sport utility vehicle'] <- 'sport utility / station wagon'
data2$VEHICLE.TYPE.CODE.4 <- tolower(data2$VEHICLE.TYPE.CODE.4)
data2$VEHICLE.TYPE.CODE.4[data2$VEHICLE.TYPE.CODE.4=='station wagon/sport utility vehicle'] <- 'sport utility / station wagon'
data2$VEHICLE.TYPE.CODE.5 <- tolower(data2$VEHICLE.TYPE.CODE.5)
data2$VEHICLE.TYPE.CODE.5[data2$VEHICLE.TYPE.CODE.5=='station wagon/sport utility vehicle'] <- 'sport utility / station wagon'

head(data2)
```
```{r}
x <- data2[!is.na(data2$CONTRIBUTING.FACTOR.VEHICLE.1),]

x$NUMBER.OF.PERSONS.F = x$NUMBER.OF.PERSONS.KILLED + x$NUMBER.OF.PERSONS.INJURED

x <- x %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% 
    summarise(NUMBER.OF.PERSONS.F = sum(NUMBER.OF.PERSONS.F),.groups = 'drop')

x <- x[order(-x$NUMBER.OF.PERSONS.F),]
head(x)
```

```{r}
x <- data2[!is.na(data2$CONTRIBUTING.FACTOR.VEHICLE.1),]

fulldata <- x %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% summarise(NUMBER.OF.PERSONS.INJURED = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')

fulldata <- fulldata[order(-fulldata$NUMBER.OF.PERSONS.INJURED),]
fulldata <- fulldata[2:20,]


brooklyn <- x %>% filter(BOROUGH == 'BROOKLYN') %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% select(CONTRIBUTING.FACTOR.VEHICLE.1, NUMBER.OF.PERSONS.INJURED) %>% summarise(NUMBER.OF.PERSONS.INJURED = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')

brooklyn <- brooklyn[order(-brooklyn$NUMBER.OF.PERSONS.INJURED),]


queens <- x %>% filter(BOROUGH == 'QUEENS') %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% select(CONTRIBUTING.FACTOR.VEHICLE.1, NUMBER.OF.PERSONS.INJURED) %>% summarise(NUMBER.OF.PERSONS.INJURED = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')

queens <- queens[order(-queens$NUMBER.OF.PERSONS.INJURED),]

manhattan <- x %>% filter(BOROUGH == 'MANHATTAN') %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% select(CONTRIBUTING.FACTOR.VEHICLE.1, NUMBER.OF.PERSONS.INJURED) %>% summarise(NUMBER.OF.PERSONS.INJURED = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')

manhattan <- manhattan[order(-manhattan$NUMBER.OF.PERSONS.INJURED),]

bronx <- x %>% filter(BOROUGH == 'BRONX') %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% select(CONTRIBUTING.FACTOR.VEHICLE.1, NUMBER.OF.PERSONS.INJURED) %>% summarise(NUMBER.OF.PERSONS.INJURED = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')

bronx <- bronx[order(-bronx$NUMBER.OF.PERSONS.INJURED),]

staten <- x %>% filter(BOROUGH == 'STATEN ISLAND') %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% select(CONTRIBUTING.FACTOR.VEHICLE.1, NUMBER.OF.PERSONS.INJURED) %>% summarise(NUMBER.OF.PERSONS.INJURED = sum(NUMBER.OF.PERSONS.INJURED),.groups = 'drop')

staten <- staten[order(-staten$NUMBER.OF.PERSONS.INJURED),]

fulldata <- merge(x=brooklyn, y=fulldata, by.x="CONTRIBUTING.FACTOR.VEHICLE.1", by.y="CONTRIBUTING.FACTOR.VEHICLE.1")
fulldata <- merge(x=queens, y=fulldata, by.x="CONTRIBUTING.FACTOR.VEHICLE.1", by.y="CONTRIBUTING.FACTOR.VEHICLE.1")
fulldata <- merge(x=manhattan, y=fulldata, by.x="CONTRIBUTING.FACTOR.VEHICLE.1", by.y="CONTRIBUTING.FACTOR.VEHICLE.1")
fulldata <- merge(x=bronx, y=fulldata, by.x="CONTRIBUTING.FACTOR.VEHICLE.1", by.y="CONTRIBUTING.FACTOR.VEHICLE.1")
fulldata <- merge(x=staten, y=fulldata, by.x="CONTRIBUTING.FACTOR.VEHICLE.1", by.y="CONTRIBUTING.FACTOR.VEHICLE.1")
names(fulldata) <- c('CONTRIBUTING.FACTOR.VEHICLE.1', '2','3','4','5','6','7')
fulldata <- data.frame(fulldata)
head(fulldata)
```

```{r}
counts <- data2 %>% drop_na(VEHICLE.TYPE.CODE.1) %>% group_by(VEHICLE.TYPE.CODE.1) %>% summarize(Count = n()) %>% arrange(desc(Count))
counts <- head(counts,20)
counts$Count <- counts$Count/1000
head(counts)
```

```{r}
x <- data2[!is.na(data2$VEHICLE.TYPE.CODE.1),]
x <- x[!is.na(x$VEHICLE.TYPE.CODE.2),]
x <- x[x$VEHICLE.TYPE.CODE.1=='ambulance',]
counts <- x %>% group_by(VEHICLE.TYPE.CODE.2) %>% summarize(Count = n()) %>% arrange(desc(Count))
counts <- head(counts,20)
counts$Count <- counts$Count
head(counts)
```

